---
title: "ChIPseq in Bioconductor exercises (part 2)"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "https://rockefelleruniversity.github.io/RU_ChIPseq/"
output: 
  html_document:
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  toMessage: false    
---

```{r, setup,include=FALSE}
toMessage <- ifelse(!is.null(params$toMessage),params$toMessage,FALSE)
message(toMessage)
```


## ChIPseq functional annotation and motif identification.

In this practical we will investigate the motifs and pathways associated with the Ebf1 transcription factor in proB-cells  and B-cells.

* Pro-B-Cell Ebf1 data has been extracted from **GSM499030**.
* B-Cell Ebf1 data has been extracted from **GSE35857**.

* **Data/ebf1_proB.bed**
* **Data/ebf1_B.bed**

1. Load the ebf1 peaks for proB and B cells into R.

```{r, include=FALSE}
library(rGREAT)
library(rtracklayer)
library(BSgenome.Mmusculus.UCSC.mm10)
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
```

```{r instalcscldcsceBSded,echo=toMessage}
suppressPackageStartupMessages(library(rtracklayer))

library(rtracklayer)
ebf1_proB <- import.bed("data/ebf1_proB.bed")
ebf1_B <- import.bed("data/ebf1_B.bed")
```

2. Find peaks in B cells which we also present in proB cells.

```{r instaaxscslldeBSccded,echo=toMessage}
ebf1_BCommon <- ebf1_B[ebf1_B %over% ebf1_proB]

```

3. Extract the sequence 100bp (+/-50bp) around the geometric centre of these peaks, write to FASTA and submit to Meme-ChIP.

```{r instacsxscsllBSded,echo=toMessage}
library(BSgenome.Mmusculus.UCSC.mm10)
toMotif <- resize(ebf1_BCommon,100,fix="center")
peaksSequences <- getSeq(BSgenome.Mmusculus.UCSC.mm10,
                         toMotif)
names(peaksSequences) <- paste0(seqnames(toMotif),":",
                                         start(toMotif),
                                         "-",
                                         end(toMotif))
writeXStringSet(peaksSequences,file="commonEBFinB.fa")

```

Results files from Meme-ChIP can be found [here](http://rockefelleruniversity.github.io/ebf1_motif/meme-chip.html)

4. Annotate common B cell peaks to genes with TSS set as +/- 500bp

```{r instacscslsxslBscsSded,echo=toMessage}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
peakAnno_Ebf1B <- annotatePeak(ebf1_BCommon, tssRegion=c(-500, 500), 
                         TxDb=TxDb.Mmusculus.UCSC.mm10.knownGene, 
                         annoDb="org.Mm.eg.db")
```

5. Create a barplot of the occurrence of differing annotations ("Promoter", "Distal Intergenic"  and "other"). 

```{r instaxxscscsllBSdsced,echo=toMessage}
myAnnotation <- as.data.frame(peakAnno_Ebf1B)
simpleAnnotation <- as.vector(myAnnotation$annotation)
simpleAnnotation[!simpleAnnotation %in% c("Promoter", "Distal Intergenic")] <- "other"
forPlot <- data.frame(Annotation=simpleAnnotation)
library(ggplot2)
ggplot(forPlot,aes(x=Annotation,fill=Annotation))+geom_bar()+coord_flip()+theme_bw()+ggtitle("Simple distribution of B-cell Ebf1 binding sites")
```

6. Perform a geneset enrichment test of genes with peaks in promoter against KEGG using goseq.

```{r instacxxsscsllBSdwced,echo=toMessage}
myAnnotation_TSS <- myAnnotation[
  myAnnotation$annotation == "Promoter",]
genesWithPeakInTSS <- unique(myAnnotation_TSS$geneId)
allGeneGR <- genes(TxDb.Mmusculus.UCSC.mm10.knownGene)
allGeneIDs <- allGeneGR$gene_id
allGenesForGOseq <- as.integer(allGeneIDs %in% genesWithPeakInTSS)
names(allGenesForGOseq) <- allGeneIDs
library(goseq)
pwf=nullp(allGenesForGOseq,"mm10","knownGene",plot.fit=FALSE)

Kegg_Ebf1Peaks <- goseq(pwf,"mm10","knownGene",
                       test.cats=c("KEGG"),
                       method="Hypergeometric")
library(KEGG.db)
xx <- as.list(KEGGPATHID2NAME)
idtoName <- cbind(names(xx),unlist(xx))
idtoName[1:5,]
KeggN_Ebf1Peaks <- merge(idtoName,Kegg_Ebf1Peaks,by=1,all=TRUE)
orderByP <- order(KeggN_Ebf1Peaks$over_represented_pvalue)
KeggN_Ebf1Peaks <- KeggN_Ebf1Peaks[orderByP,]
KeggN_Ebf1Peaks[1:5,]
```

7. Identify Ebf1 peaks  only in Bcells and Ebf1 peaks only in proB cells.


```{r instascxsxslldeBSccded,echo=toMessage}
ebf1_Bunique<- ebf1_B[!ebf1_B %over% ebf1_proB]
ebf1_proBunique<- ebf1_proB[!ebf1_proB %over% ebf1_B]

```

8 Using rGREAT identify the MSigDB Pathways enriched in Ebf1 peaks  only in Bcells and Ebf1 peaks only in proB cells.


```{r v,echo=toMessage}
library(rGREAT)
great_Job <- submitGreatJob(ebf1_Bunique,species="mm10",request_interval=1,version = "3.0.0")
availableCategories(great_Job)
great_ResultTable = getEnrichmentTables(great_Job,
                                        category="Pathway Data")
names(great_ResultTable)
msigPath_Bcells <- great_ResultTable[["MSigDB Pathway"]]

msigPath_Bcells[1:10,]

great_Job <- submitGreatJob(ebf1_proBunique,species="mm10",request_interval=1,version = "3.0.0")
availableCategories(great_Job)
great_ResultTable = getEnrichmentTables(great_Job,
                                        category="Pathway Data")
names(great_ResultTable)
msigPath_proBcells <- great_ResultTable[["MSigDB Pathway"]]


msigPath_proBcells[1:10,]
```
